Empirical Bayes analysis for detection of gene heterosis in RNAseq data

### 2014-12-17 ebayes_rnaseq_heterosis.tex
through results

@@

\begin{figure}
\centerline{\includegraphics[width=\textwidth]{normality}}
\caption{One and two-dimensional histograms for gene-specific parameters from a single MCMC iteration, $\theta_g^{(m)}$,  for the B73 $\times$ Mo17 maize experiment. For reference, the one dimensional histograms have the marginal probability density functions from $N(\hat{\eta},\hat{\Sigma})$.}
\label{f:normality}
\end{figure}

@@

Put your final comments here.

>>>>

% Compare independence and volcano plots

@@

With almost 30,000 genes in our data set, we can evaluate the normality assumption for the gene-specific parameters. Figure \ref{f:normality} provides histograms of components of $\theta$ from a single MCMC iteration. If the prior assumption of normality is correct, we would expect these histograms to show approximate normality. The marginal histograms show a bimodal distribution for $\mu$ with a secondary peak around relatively small values. The bivariate histograms amongst the components of $\mu$ show a high correlation but with a heavy-tail toward small values of both components of $\mu$. With stricter gene-inclusion criterion, it is possible both the secondary peak and the heavy-tailedness could be eliminated. The histogram for $\log(\psi)$ displays a heavy-tail to the right and the bivariate histograms show a pattern of near independence between $\log(\psi)$ and $\mu_g$ until $\mu_g$ passes a threshold at which point $\log(\psi)$ appears to decrease. Both of these phenomenon are consistent with observations in the literature that the negative-binomial overdispersion parameter decreases as the mean increases, but the bivariate histograms appear to indicate a nonlinear relationship between the means and the log of the overdispersion parameter.

\begin{figure}
\centerline{\includegraphics[width=\textwidth]{normality}}
\caption{Marginal and bivariate histograms for gene-specific parameters from a single MCMC iteration, $\theta_g^{(m)}$,  for the B73 $\times$ Mo17 maize experiment. For reference, the marginal histograms have a dashed line indicating the appropriate marginal probability density functions from $N(\hat{\eta},\hat{\Sigma})$.}
\label{f:normality}
\end{figure}

Marginal and bivariate histograms for the model with an estimated covariance look similar although the estimated distribution $N(\hat{\eta},\hat{\Sigma})$ has increased variance for $\mu$ and decreased variance for $\log(\psi)$ as was previously seen in Table \ref{t:map} (results not shown).

@@
